### 2015-06-19 reports/paper/prior_movement.tex

Kalman filter uses observed variables to infer hidden variables and use them to help predict the next state. In our system, observed variables $z$ are the $(x,y)$ coordinates of the localized acoustic source:
\[
z = \left[\begin{array}{c}
x\\
y\\
\end{array}\right]
\]
We can also model unobserved motion variables such as velocity $(\dot{x}, \dot{y})$ and acceleration $(\ddot{x}, \ddot{y})$. Then the state variables we are tracking $\mathbf{x}$ can be represented as:
\[
\mathbf{x} = \left[x, y, \dot{x}, \dot{y}, \ddot{x}, \ddot{y}\right]^T
\]
Note that by modeling up to acceleration, we are implicitly assuming higher order motion variables are constant (such as jerk). Kalman filter includes a process noise term $Q$ that can be used to partially reduce this effect, but there will exist systematic tracking bias there is acceleration change.

At any time instant, Kalman filter can use current state information to infer the predicted next state $\mathbf{x}^-$ and the predicted uncertainty $P^-$:
\begin{eqnarray}
\mathbf{x}^- & = &F \mathbf{x} + B_ku_k\\
P^- & = & FPF^T + Q\\
\end{eqnarray}
where
\begin{itemize}
\item F is the state transition matrix. In our system, the next coordinate output can be computed with laws of physics using current position, velocity and acceleration:
\[
F  =  \left(\begin{array}{cccccc}
1 & 0 & \frac{1}{2} & 0 & \frac{1}{2} & 0\\
1 & 0 & \frac{1}{2} & 0 & \frac{1}{2} & 0 \\
\end{array}\right)
\]
\item $F_k \hat{x}_k + B_k u_k$ describes the station transition form time $k$ to $k+1$. The Kalman filter is a general framework which allows not only tracking but also controlling the object. I these situations $u_k$ is the control input and $B$ describes the control input model. For our application, we are only tracking and not controlling the movement, $B_ku_k = 0$.
\end{itemize}
and an update step:
\begin{eqnarray}
y_k & = & z_k - H_k\hat{x}_k^-\\
S_k & = & H_kP_k^-H_k^T + R_k \\
K_k & = & P_k^-H_k^TS_k^{-1}\\
\hat{x}_k & = & \hat{x}_k^- + K_ky\\
P_k & = & (I-K_kH_k)P_k^-\\
\end{eqnarray}
where:
\begin{itemize}
\item $x$ is the state variable
\item $P$ is the covariance on the state variable $x$
\item $z$ is the measurement.
\item $y$ is the residue in the measurement sapce. 
\item $H$ is the measurement function that transforms from state space to measurement space.
\item $R$ is the measurement noise matrix that models sensor noise as a covariance matrix
\item $Q$ models the process noise
\end{itemize}

In our system,

>>>>

Kalman filter uses observed variables to infer hidden variables and use them to help predict the next state. In our system, observed variables $z$ are the $(x,y)$ coordinates of the localized acoustic source:
\[
z = \left[\begin{array}{c}
x\\
y\\
\end{array}\right]
\]
We can also model unobserved motion variables such as velocity $(\dot{x}, \dot{y})$ and acceleration $(\ddot{x}, \ddot{y})$. Then the state variables we are tracking $\mathbf{x}$ can be represented as:
\[
\mathbf{x} = \left[x, y, \dot{x}, \dot{y}, \ddot{x}, \ddot{y}\right]^T
\]
Internally, the Kalman filter also keeps track of the uncertainty on the state variables. It is represented as a covariance matrix $P$ on state variables.

Note that by modeling up to acceleration, we are implicitly assuming higher order motion variables are constant (such as jerk). Kalman filter includes a process noise term $Q$ that can be used to partially reduce this effect, but there will exist systematic tracking bias there is acceleration change.

At any time instant, Kalman filter can use current state information to infer the predicted next state $\mathbf{x}^-$ and the predicted uncertainty $P^-$:
\begin{eqnarray}
\mathbf{x}^- & = &F \mathbf{x} + Bu\\
P^- & = & FPF^T + Q\\
\end{eqnarray}
where
\begin{itemize}
\item F is the state transition matrix. In our system, next coordinates $(x,y)$ can be computed with laws of physics using current position, velocity and acceleration:
\[
F  =  \left(\begin{array}{cccccc}
1 & 0 & \delta t & 0 & \frac{1}{2}\delta t^2 & 0 \\
0 & 1 & 0 & \delta t & 0 & \frac{1}{2} \delta t^2 \\
0 & 0 & 1 & 0 & \delta t & 0\\
0 & 0 & 0 & 1 & 0 & \delta t\\
0 & 0 & 0 & 0 & 1 & 0 \\
0 & 0 & 0 & 0 & 0 & 1 \\
\end{array}\right)
\], where $\delta t$ represents the time difference between current time and last state update. 
\item The Kalman filter is a general framework which allows not only tracking but also controlling the object. In these situations $u$ is the control input and $B$ describes the control input model. For our application, we are only tracking and not controlling the movement, $Bu = 0$.
\item $Q$ models the process noise
\end{itemize}

After the observation for the next coordinates are made, then the Kalman filter calculates the residue between the prediction $\mathbf{x}^-$ and the measurement $z$:
state $z$:
\[
y = z - H\mathbf{x}^-
\]
where $H$ is the measurement function that transforms from state space to measurement space. In our example, the transformation is simply taking the location $(x,y)$ from the state space:
\[
H = \left(\begin{array}{cccccc}
1 & 0 & 0 & 0 & 0 & 0 \\
0 & 1 & 0 & 0 & 0 & 0 \\
\end{array}\right)
\]

Then the Kalman filter updates the states estimates $\mathbf{x}$ and states uncertainty estimates $P$:
\begin{eqnarray}
\mathbf{x} & = & \mathbf{x}^- + Ky\\
P & = & (I-KH)P^-
\end{eqnarray}
where $K$ is the Kalman gain:
\begin{eqnarray}
K & = & P^-H^TS^{-1}\\
S & = & HP^-H^T + R
\end{eqnarray}
$R$ is the measurement noise matrix that models the localization system's output noise as a covariance matrix.
